MISUSES AND INACCURACIES OF PAGEL'S LAMBDA ML ESTIMATION METHODS

### 2020-04-20 Manuscript/Manuscript-1-DCA.Rmd
new edits

Several studies have investigated the statistical properties of methods for estimating phylogenetic signal under various conditions [e.g., @Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @MolinaVenegas2017; see also @Revell_et_al2008; @Revell2010]. These have largely focused on the ability of methods to detect the presence of phylogenetic signal (i.e., type I and type II error rates) under complex models of evolutionary change, across a range of phylogeny sizes, and with varying degrees of phylogenetic uncertainty or unresolved topologies. In terms of parameter estimation, Revell [-@Revell2010] found that regression parameters were accurately estimated when $\lambda$ was included during phylogenetic regression, and Munkmuller et al. [-@Munkemuller_et_al2012] demonstrated that estimates of phylogenetic signal obtained using various measures generally increased when input levels of phylogenetic signal were stronger. However, the precision of those estimates could not be determined, because the input levels of phylogenetic signal were simulated via a scaling factor ($w$) that was not directly comparable to the measures of phylogenetic signal being compared [see @Munkemuller_et_al2012]. One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the lack of precision in parameter estimation. However, this conclusion assumes that the precision of parameter estimation remains constant across the range of values ($\lambda = 0 \to 1$); an assumption that to date, has not been verified. Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.  \hfill\break

  In this study, we evaluate the precision of Pagel's $\lambda$ to estimate known levels of phylogenetic signal in phenotypic data. First we use computer simulations across differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances. For comparison, we also evaluate estimates of $Kappa$ and its empirically-derived effect size ($Z_K$) for this purpose. Additionally, we use simulations to determine how the inclusion of $\lambda$ in phylogenetic regression and ANOVA (i.e., PGLS) affects parameter estimation, and whether levels of phylogenetic signal estimated in a PGLS framework are accurate. We then survey the recent macroevolutionary literature for published papers containing estimates of $\lambda$ from empirical datasets, and compare these empirical estimates to patterns gleaned from our computer simulations. In general we find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not. Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength. Alternatively, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more uniform across the range of input values, and estimates of phylogenetic signal with this measure increase in a uniform manner with increasing phylogenetic signal. We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure should be treated with caution. By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and comparing the strength of phylogenetic signal across datasets.

@@

We conducted a series of computer simulations to evaluate the ability of Pagel's [-@Pagel1999] $\lambda$ to estimate known levels of phylogenetic signal. Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine the extent to which tree shape affected our findings (see Supplemental Material). Our first set of simulations evaluated the extent to which values of $\lambda$ estimated from the data corresponded with actual input levels of phylogenetic signal. For these simulations we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$). Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda$. We used a set of values that encompassed the entire range of $\lambda$: ($\lambda_{in} = 0.0 - 1.0$; in 21 intervals of 0.05 units), resulting in 1050 scaled phylogenies at each level of species richness ($n$). Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with known and differing levels of phylogenetic signal. These varied from datasets with no phylogenetic signal (when $\lambda_{in} =0$) to datsets with phylogenetic signal corresponding to what was expected under Brownian motion (when $\lambda_{in} =1$). \hfill\break

Using the simulated phylogenies we estimated the phylogenetic signal in each dataset using Pagel's $\lambda$. Next, for the 1050 datasets at each level of species richness ($n$), we assessed the relationship between $\lambda_{in}$ and $\lambda_{est}$ using regression. Additionally, the precision of $\lambda_{est}$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$). Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset. Here, the effect size was estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution [for discussion see: @AdamsCollyer2019]. Because $Kappa$ and $Z_K$ differ in scale from $\lambda$, both measures were linearly normalized to a standard uniform distribution ($0\rightarrow1$), and the variation in the set of normalized $Kappa$ and $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were estimated for comparison. \hfill\break

@@

\textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips. \hfill\break

\textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases. However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern. \hfill\break

\textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

@@

\singlespacing \textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. As trees increase in size, the estimates more closely resemble the input lambdas, however considerable and concerning variation is apparent in trees smaller than those with 200 tips. \hfill\break

@@

\singlespacing \textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes, the mean estimated slope matches the input slope, and as trees increase in size, the variance around this mean estimate decreases. However, for trees with fewer than 200 tips, the error around the estimated slope is considerable, where these analyses frequently estimate slopes in the opposite direction of the known pattern. \hfill\break

@@

\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

>>>>

Several studies have investigated the statistical properties of methods for estimating phylogenetic signal under various conditions [e.g., @Munkemuller_et_al2012; @Pavoine2012;  @DinizFilho2012; @MolinaVenegas2017; see also @Revell_et_al2008; @Revell2010]. These have largely focused on the ability of methods to detect the presence of phylogenetic signal (i.e., type I and type II error rates) under complex models of evolutionary change, across a range of phylogeny sizes, and with varying degrees of phylogenetic uncertainty or unresolved topologies. In terms of parameter estimation, Revell [-@Revell2010] found that regression parameters were accurately estimated when $\lambda$ was included during phylogenetic regression, and Munkmuller et al. [-@Munkemuller_et_al2012] demonstrated that estimates of phylogenetic signal obtained using various measures generally increased when input levels of phylogenetic signal were stronger. However, the precision of those estimates could not be determined, because the input levels of phylogenetic signal were simulated via a scaling factor ($w$) that was not directly comparable to the measures of phylogenetic signal being compared [see @Munkemuller_et_al2012]. One study [@Boettiger_et_al2012] found that estimates of Pagel's $\lambda$ displayed less variation when data were simulated on a large phylogeny ($N=281$) as compared to a small one ($N=13$), concluding that insufficient data (i.e., the number of species) was the underlying cause of the lack of precision in parameter estimation. However, this conclusion assumes that the precision in estimating $\lambda$ remains constant across its range ($\lambda = 0 \to 1$); an assumption that to date, has not been verified. Thus, despite widespread use of Pagel's [-@Pagel1999] $\lambda$ in macroevolutionary studies, at present, we still lack a general understanding of the precision with which $\lambda$ can estimate levels of phylogenetic signal in phenotypic datasets.  \hfill\break

  In this study, we evaluate the precision of Pagel's $\lambda$ in estimating known levels of phylogenetic signal in phenotypic data. First we use computer simulations across differing numbers of species, differently shaped phylogenies, and differing input levels of phylogenetic signal, to explore the degree to which $\lambda$ correctly identifies known levels of phylogenetic signal, and under what circumstances. For comparison, we also evaluate estimates of $Kappa$ and its empirically-derived effect size ($Z_K$) for this purpose. Additionally, we use simulations to determine how the inclusion of $\lambda$ in phylogenetic regression and ANOVA (i.e., PGLS) affects parameter estimation, and whether levels of phylogenetic signal estimated in a PGLS framework are accurate. We then survey the recent macroevolutionary literature for published papers containing estimates of $\lambda$ from empirical datasets, and compare these empirical estimates to patterns gleaned from our computer simulations. In general we find that while PGLS parameters (e.g., $\beta$) are accurately estimated with the inclusion of phylogenetic signal, estimates of $\lambda$ are not. Further, we find that estimates of $\lambda$ vary widely for a given input value of phylogenetic signal, and that this variation is not constant across the range of input signal, with decreased precision when phylogenetic signal is of intermediate strength. By contrast, variation across effect sizes ($Z_K$) obtained from $Kappa$ is far more consistent across the range of input values, and that the precision of estimating $Z_K$ is preferable to that observed using $\lambda$. We conclude that estimates of phylogenetic signal using Pagel's $\lambda$ are often innacurate, and thus interpreting strength of phylogenetic signal in phenotypic datasets based on this measure should be treated with caution. By contrast, effect sizes obtained from $Kappa$ hold promise for characterizing phylogenetic signal, and comparing the strength of phylogenetic signal across datasets.

@@

We conducted a series of computer simulations to evaluate the ability of Pagel's [-@Pagel1999] $\lambda$ to estimate known levels of phylogenetic signal. Our primary simulations were based on pure-birth phylogenies; however, we also evaluated patterns on both balanced and pectinate trees to determine whether tree shape affected our findings (see Supporting Information). Our first set of simulations evaluated the extent to which values of $\lambda$ estimated from the data corresponded with actual input levels of phylogenetic signal. For these simulations we generated 50 pure-birth phylogenies at each of six different tree sizes, ranging from 32 to 1024 taxa ($n=2^5 - 2^{10}$). Next, we rescaled the simulated phylogenies by multiplying the internal branches by $\lambda$. We used a set of values that encompassed the entire range of $\lambda$: ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units), resulting in 1050 scaled phylogenies at each level of species richness ($n$). Continuous traits were then simulated on each phylogeny under a Brownian motion model of evolution to obtain datasets with known and differing levels of phylogenetic signal. These varied from datasets with no phylogenetic signal (when $\lambda_{in} =0$) to datsets with phylogenetic signal corresponding to what was expected under Brownian motion (when $\lambda_{in} =1$). \hfill\break

Using the simulated phylogenies we estimated the phylogenetic signal in each dataset using Pagel's $\lambda$. Next, for the 1050 datasets at each level of species richness ($n$), we assessed the relationship between $\lambda_{in}$ and $\lambda_{est}$ using regression. Additionally, the precision of $\lambda_{est}$ was approximated by observing the variation of $\lambda_{est}$ obtained at each input level of phylogenetic signal ($\lambda_{in}$). Finally, for comparison we estimated the degree of phylogenetic signal in each dataset using $Kappa$ [@Blomberg_et_al2003; @Adams2014a], and from this calculated its effect size [$Z_K$: sensu @AdamsCollyer2016; @AdamsCollyer2019b] for each dataset. Here, the effect size was estimated as: $Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}$, where $K_{obs}$ was the observed $Kappa$, and $\mu_K$ and $\sigma_K$ were the mean and standard deviation of the empirical sampling distribution of values obtained from the permutation distribution [for discussion see: @AdamsCollyer2019]. Because $Z_K$ differs in scale from $\lambda$, we used a linear normalization to standardize $Z_K$ to a uniform distribution ($0\rightarrow1$), and variation in the set of normalized $Z_K$ values at each input level of phylogenetic signal ($\lambda_{in}$) were then estimated for comparison. \hfill\break

@@

\textbf{Figure 1}. Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 2}. Variation in estimates of phylogenetic signal across input levels of phylogenetic signal. **A**: Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), **B** Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), **C** Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species. **D** Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.  \hfill\break

\textbf{Figure 3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($|Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 4}. Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

@@

\singlespacing \textbf{Figure 1}. Precision of Pagel's $\lambda$ across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.

@@

\singlespacing \textbf{Figure 2}.  Variation in estimates of phylogenetic signal across input levels of phylogenetic signal. **A**: Estimates of Pagel's $\lambda$ for data simulated on phylogenies with 128 taxa ($n=128$), **B** Estimates of $Z_K$ for data simulated on phylogenies with 128 taxa ($n=128$), **C** Variance in the variation of $\lambda_{est}$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species. **D** Variance in the variation of $Z_K$ across input levels of phylogenetic signal, estimated on phylogenies containing differing numbers of species.

@@

\singlespacing \textbf{Figure 3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic regression ($|Y\sim X$), across known levels of input phylogenetic signal ($\lambda_{in}$) on phylogenies of various sizes. As phylogenies increase in size, variation in $\lambda_{in}$ decreases; however the precision is not constant across the range of input levels ($\lambda_{in}: 0 \to 1$), and is highest at intermediate levels of phylogenetic signal.

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig4.png")

\singlespacing \textbf{Figure 4}.Frequency of estimated lambda values published in manuscripts in 2019. The majority of these values were close to 0 or 1, and from phylogenies with fewer than 200 taxa.

### 2020-04-20 Manuscript/Manuscript-1-DCA.tex
new edits

Several studies have investigated the statistical properties of methods
for estimating phylogenetic signal under various conditions (e.g.,
Munkemuller et al. 2012; Pavoine and Ricotta 2012; Diniz-Filho et al.
2012; Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008;
Revell 2010). These have largely focused on the ability of methods to
detect the presence of phylogenetic signal (i.e., type I and type II
error rates) under complex models of evolutionary change, across a range
of phylogeny sizes, and with varying degrees of phylogenetic uncertainty
or unresolved topologies. In terms of parameter estimation, Revell
(2010) found that regression parameters were accurately estimated when
\(\lambda\) was included during phylogenetic regression, and Munkmuller
et al.~(2012) demonstrated that estimates of phylogenetic signal
obtained using various measures generally increased when input levels of
phylogenetic signal were stronger. However, the precision of those
estimates could not be determined, because the input levels of
phylogenetic signal were simulated via a scaling factor (\(w\)) that was
not directly comparable to the measures of phylogenetic signal being
compared (see Munkemuller et al. 2012). One study (Boettiger et al.
2012) found that estimates of Pagel's \(\lambda\) displayed less
variation when data were simulated on a large phylogeny (\(N=281\)) as
compared to a small one (\(N=13\)), concluding that insufficient data
(i.e., the number of species) was the underlying cause of the lack of
precision in parameter estimation. However, this conclusion assumes that
the precision of parameter estimation remains constant across the range
of values (\(\lambda = 0 \to 1\)); an assumption that to date, has not
been verified. Thus, despite widespread use of Pagel's (1999)
\(\lambda\) in macroevolutionary studies, at present, we still lack a
general understanding of the precision with which \(\lambda\) can
estimate levels of phylogenetic signal in phenotypic datasets.
\hfill\break

In this study, we evaluate the precision of Pagel's \(\lambda\) to
estimate known levels of phylogenetic signal in phenotypic data. First
we use computer simulations across differing numbers of species,
differently shaped phylogenies, and differing input levels of
phylogenetic signal, to explore the degree to which \(\lambda\)
correctly identifies known levels of phylogenetic signal, and under what
circumstances. For comparison, we also evaluate estimates of \(Kappa\)
and its empirically-derived effect size (\(Z_K\)) for this purpose.
Additionally, we use simulations to determine how the inclusion of
\(\lambda\) in phylogenetic regression and ANOVA (i.e., PGLS) affects
parameter estimation, and whether levels of phylogenetic signal
estimated in a PGLS framework are accurate. We then survey the recent
macroevolutionary literature for published papers containing estimates
of \(\lambda\) from empirical datasets, and compare these empirical
estimates to patterns gleaned from our computer simulations. In general
we find that while PGLS parameters (e.g., \(\beta\)) are accurately
estimated with the inclusion of phylogenetic signal, estimates of
\(\lambda\) are not. Further, we find that estimates of \(\lambda\) vary
widely for a given input value of phylogenetic signal, and that this
variation is not constant across the range of input signal, with
decreased precision when phylogenetic signal is of intermediate
strength. Alternatively, variation across effect sizes (\(Z_K\))
obtained from \(Kappa\) is far more uniform across the range of input
values, and estimates of phylogenetic signal with this measure increase
in a uniform manner with increasing phylogenetic signal. We conclude
that estimates of phylogenetic signal using Pagel's \(\lambda\) are
often innacurate, and thus interpreting strength of phylogenetic signal
in phenotypic datasets based on this measure should be treated with
caution. By contrast, effect sizes obtained from \(Kappa\) hold promise
for characterizing phylogenetic signal, and comparing the strength of
phylogenetic signal across datasets.

@@

We conducted a series of computer simulations to evaluate the ability of
Pagel's (1999) \(\lambda\) to estimate known levels of phylogenetic
signal. Our primary simulations were based on pure-birth phylogenies;
however, we also evaluated patterns on both balanced and pectinate trees
to determine the extent to which tree shape affected our findings (see
Supplemental Material). Our first set of simulations evaluated the
extent to which values of \(\lambda\) estimated from the data
corresponded with actual input levels of phylogenetic signal. For these
simulations we generated 50 pure-birth phylogenies at each of six
different tree sizes, ranging from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)).
Next, we rescaled the simulated phylogenies by multiplying the internal
branches by \(\lambda\). We used a set of values that encompassed the
entire range of \(\lambda\): (\(\lambda_{in} = 0.0 - 1.0\); in 21
intervals of 0.05 units), resulting in 1050 scaled phylogenies at each
level of species richness (\(n\)). Continuous traits were then simulated
on each phylogeny under a Brownian motion model of evolution to obtain
datasets with known and differing levels of phylogenetic signal. These
varied from datasets with no phylogenetic signal (when
\(\lambda_{in} =0\)) to datsets with phylogenetic signal corresponding
to what was expected under Brownian motion (when \(\lambda_{in} =1\)).
\hfill\break

Using the simulated phylogenies we estimated the phylogenetic signal in
each dataset using Pagel's \(\lambda\). Next, for the 1050 datasets at
each level of species richness (\(n\)), we assessed the relationship
between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
Additionally, the precision of \(\lambda_{est}\) was approximated by
observing the variation of \(\lambda_{est}\) obtained at each input
level of phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison
we estimated the degree of phylogenetic signal in each dataset using
\(Kappa\) (Blomberg et al. 2003; Adams 2014a), and calculated its effect
size (\(Z_K\): sensu Adams and Collyer 2016, 2019a) for each dataset.
Here, the effect size was estimated as:
\(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean and
standard deviation of the empirical sampling distribution of values
obtained from the permutation distribution (for discussion see: Adams
and Collyer 2019b). Because \(Kappa\) and \(Z_K\) differ in scale from
\(\lambda\), both measures were linearly normalized to a standard
uniform distribution (\(0\rightarrow1\)), and the variation in the set
of normalized \(Kappa\) and \(Z_K\) values at each input level of
phylogenetic signal (\(\lambda_{in}\)) were estimated for comparison.
\hfill\break

@@

\textbf{Figure 1}. Accuracy of Pagel's lambda estimations across known
lambda inputs on various tree sizes. As trees increase in size, the
estimates more closely resemble the input lambdas, however considerable
and concerning variation is apparent in trees smaller than those with
200 tips. \hfill\break

\textbf{Figure 2}. Estimated ANOVA slopes under PGLS. Across tree sizes,
the mean estimated slope matches the input slope, and as trees increase
in size, the variance around this mean estimate decreases. However, for
trees with fewer than 200 tips, the error around the estimated slope is
considerable, where these analyses frequently estimate slopes in the
opposite direction of the known pattern. \hfill\break

\textbf{Figure 3}. Frequency of estimated lambda values published in
manuscripts in 2019. The majority of these values were close to 0 or 1,
and from phylogenies with fewer than 200 taxa.

@@

\singlespacing \textbf{Figure 1}. Accuracy of Pagel's lambda estimations
across known lambda inputs on various tree sizes. As trees increase in
size, the estimates more closely resemble the input lambdas, however
considerable and concerning variation is apparent in trees smaller than
those with 200 tips. \hfill\break

@@

\singlespacing \textbf{Figure 2}. Estimated ANOVA slopes under PGLS.
Across tree sizes, the mean estimated slope matches the input slope, and
as trees increase in size, the variance around this mean estimate
decreases. However, for trees with fewer than 200 tips, the error around
the estimated slope is considerable, where these analyses frequently
estimate slopes in the opposite direction of the known pattern.
\hfill\break

@@

\includegraphics[width=0.95\linewidth]{Fig3}

\singlespacing \textbf{Figure 3}. Frequency of estimated lambda values
published in manuscripts in 2019. The majority of these values were
close to 0 or 1, and from phylogenies with fewer than 200 taxa.

>>>>

Several studies have investigated the statistical properties of methods
for estimating phylogenetic signal under various conditions (e.g.,
Munkemuller et al. 2012; Pavoine and Ricotta 2012; Diniz-Filho et al.
2012; Molina-Venegas and Rodriguez 2017; see also Revell et al. 2008;
Revell 2010). These have largely focused on the ability of methods to
detect the presence of phylogenetic signal (i.e., type I and type II
error rates) under complex models of evolutionary change, across a range
of phylogeny sizes, and with varying degrees of phylogenetic uncertainty
or unresolved topologies. In terms of parameter estimation, Revell
(2010) found that regression parameters were accurately estimated when
\(\lambda\) was included during phylogenetic regression, and Munkmuller
et al.~(2012) demonstrated that estimates of phylogenetic signal
obtained using various measures generally increased when input levels of
phylogenetic signal were stronger. However, the precision of those
estimates could not be determined, because the input levels of
phylogenetic signal were simulated via a scaling factor (\(w\)) that was
not directly comparable to the measures of phylogenetic signal being
compared (see Munkemuller et al. 2012). One study (Boettiger et al.
2012) found that estimates of Pagel's \(\lambda\) displayed less
variation when data were simulated on a large phylogeny (\(N=281\)) as
compared to a small one (\(N=13\)), concluding that insufficient data
(i.e., the number of species) was the underlying cause of the lack of
precision in parameter estimation. However, this conclusion assumes that
the precision in estimating \(\lambda\) remains constant across its
range (\(\lambda = 0 \to 1\)); an assumption that to date, has not been
verified. Thus, despite widespread use of Pagel's (1999) \(\lambda\) in
macroevolutionary studies, at present, we still lack a general
understanding of the precision with which \(\lambda\) can estimate
levels of phylogenetic signal in phenotypic datasets. \hfill\break

In this study, we evaluate the precision of Pagel's \(\lambda\) in
estimating known levels of phylogenetic signal in phenotypic data. First
we use computer simulations across differing numbers of species,
differently shaped phylogenies, and differing input levels of
phylogenetic signal, to explore the degree to which \(\lambda\)
correctly identifies known levels of phylogenetic signal, and under what
circumstances. For comparison, we also evaluate estimates of \(Kappa\)
and its empirically-derived effect size (\(Z_K\)) for this purpose.
Additionally, we use simulations to determine how the inclusion of
\(\lambda\) in phylogenetic regression and ANOVA (i.e., PGLS) affects
parameter estimation, and whether levels of phylogenetic signal
estimated in a PGLS framework are accurate. We then survey the recent
macroevolutionary literature for published papers containing estimates
of \(\lambda\) from empirical datasets, and compare these empirical
estimates to patterns gleaned from our computer simulations. In general
we find that while PGLS parameters (e.g., \(\beta\)) are accurately
estimated with the inclusion of phylogenetic signal, estimates of
\(\lambda\) are not. Further, we find that estimates of \(\lambda\) vary
widely for a given input value of phylogenetic signal, and that this
variation is not constant across the range of input signal, with
decreased precision when phylogenetic signal is of intermediate
strength. By contrast, variation across effect sizes (\(Z_K\)) obtained
from \(Kappa\) is far more consistent across the range of input values,
and that the precision of estimating \(Z_K\) is preferable to that
observed using \(\lambda\). We conclude that estimates of phylogenetic
signal using Pagel's \(\lambda\) are often innacurate, and thus
interpreting strength of phylogenetic signal in phenotypic datasets
based on this measure should be treated with caution. By contrast,
effect sizes obtained from \(Kappa\) hold promise for characterizing
phylogenetic signal, and comparing the strength of phylogenetic signal
across datasets.

@@

We conducted a series of computer simulations to evaluate the ability of
Pagel's (1999) \(\lambda\) to estimate known levels of phylogenetic
signal. Our primary simulations were based on pure-birth phylogenies;
however, we also evaluated patterns on both balanced and pectinate trees
to determine whether tree shape affected our findings (see Supporting
Information). Our first set of simulations evaluated the extent to which
values of \(\lambda\) estimated from the data corresponded with actual
input levels of phylogenetic signal. For these simulations we generated
50 pure-birth phylogenies at each of six different tree sizes, ranging
from 32 to 1024 taxa (\(n=2^5 - 2^{10}\)). Next, we rescaled the
simulated phylogenies by multiplying the internal branches by
\(\lambda\). We used a set of values that encompassed the entire range
of \(\lambda\): (\(\lambda_{in} = 0.0 \to 1.0\); in 21 intervals of 0.05
units), resulting in 1050 scaled phylogenies at each level of species
richness (\(n\)). Continuous traits were then simulated on each
phylogeny under a Brownian motion model of evolution to obtain datasets
with known and differing levels of phylogenetic signal. These varied
from datasets with no phylogenetic signal (when \(\lambda_{in} =0\)) to
datsets with phylogenetic signal corresponding to what was expected
under Brownian motion (when \(\lambda_{in} =1\)). \hfill\break

Using the simulated phylogenies we estimated the phylogenetic signal in
each dataset using Pagel's \(\lambda\). Next, for the 1050 datasets at
each level of species richness (\(n\)), we assessed the relationship
between \(\lambda_{in}\) and \(\lambda_{est}\) using regression.
Additionally, the precision of \(\lambda_{est}\) was approximated by
observing the variation of \(\lambda_{est}\) obtained at each input
level of phylogenetic signal (\(\lambda_{in}\)). Finally, for comparison
we estimated the degree of phylogenetic signal in each dataset using
\(Kappa\) (Blomberg et al. 2003; Adams 2014a), and from this calculated
its effect size (\(Z_K\): sensu Adams and Collyer 2016, 2019a) for each
dataset. Here, the effect size was estimated as:
\(Z_K=\frac{K_{obs}-\mu_K}{\sigma_K}\), where \(K_{obs}\) was the
observed \(Kappa\), and \(\mu_K\) and \(\sigma_K\) were the mean and
standard deviation of the empirical sampling distribution of values
obtained from the permutation distribution (for discussion see: Adams
and Collyer 2019b). Because \(Z_K\) differs in scale from \(\lambda\),
we used a linear normalization to standardize \(Z_K\) to a uniform
distribution (\(0\rightarrow1\)), and variation in the set of normalized
\(Z_K\) values at each input level of phylogenetic signal
(\(\lambda_{in}\)) were then estimated for comparison. \hfill\break

@@

\textbf{Figure 1}. Precision of Pagel's \(\lambda\) across known levels
of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
various sizes. As phylogenies increase in size, variation in
\(\lambda_{in}\) decreases; however the precision is not constant across
the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 2}. Variation in estimates of phylogenetic signal across
input levels of phylogenetic signal. \textbf{A}: Estimates of Pagel's
\(\lambda\) for data simulated on phylogenies with 128 taxa (\(n=128\)),
\textbf{B} Estimates of \(Z_K\) for data simulated on phylogenies with
128 taxa (\(n=128\)), \textbf{C} Variance in the variation of
\(\lambda_{est}\) across input levels of phylogenetic signal, estimated
on phylogenies containing differing numbers of species. \textbf{D}
Variance in the variation of \(Z_K\) across input levels of phylogenetic
signal, estimated on phylogenies containing differing numbers of
species. \hfill\break

\textbf{Figure 3}. Precision of Pagel's \(\lambda\) when incorporated in
phylogenetic regression (\(|Y\sim X\)), across known levels of input
phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
As phylogenies increase in size, variation in \(\lambda_{in}\)
decreases; however the precision is not constant across the range of
input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at intermediate
levels of phylogenetic signal. \hfill\break

\textbf{Figure 4}. Frequency of estimated lambda values published in
manuscripts in 2019. The majority of these values were close to 0 or 1,
and from phylogenies with fewer than 200 taxa.

@@

\singlespacing \textbf{Figure 1}. Precision of Pagel's \(\lambda\)
across known levels of input phylogenetic signal (\(\lambda_{in}\)) on
phylogenies of various sizes. As phylogenies increase in size, variation
in \(\lambda_{in}\) decreases; however the precision is not constant
across the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is
highest at intermediate levels of phylogenetic signal.

@@

\singlespacing \textbf{Figure 2}. Variation in estimates of phylogenetic
signal across input levels of phylogenetic signal. \textbf{A}: Estimates
of Pagel's \(\lambda\) for data simulated on phylogenies with 128 taxa
(\(n=128\)), \textbf{B} Estimates of \(Z_K\) for data simulated on
phylogenies with 128 taxa (\(n=128\)), \textbf{C} Variance in the
variation of \(\lambda_{est}\) across input levels of phylogenetic
signal, estimated on phylogenies containing differing numbers of
species. \textbf{D} Variance in the variation of \(Z_K\) across input
levels of phylogenetic signal, estimated on phylogenies containing
differing numbers of species.

@@

\singlespacing \textbf{Figure 3}. Precision of Pagel's \(\lambda\) when
incorporated in phylogenetic regression (\(|Y\sim X\)), across known
levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
various sizes. As phylogenies increase in size, variation in
\(\lambda_{in}\) decreases; however the precision is not constant across
the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
intermediate levels of phylogenetic signal.

\includegraphics[width=0.95\linewidth]{Fig4}

\singlespacing \textbf{Figure 4}.Frequency of estimated lambda values
published in manuscripts in 2019. The majority of these values were
close to 0 or 1, and from phylogenies with fewer than 200 taxa.
